---
title: 'DATA 604 Assignment 8: Input Analysis'
author: "Dan Smilowitz"
date: "March 25, 2017"
output:
  pdf_document: 
    fig_height: 4.25
    fig_caption: no
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.align='center', comment=NA)
```

```{r libraries-hidden, echo=FALSE}
library(fitdistrplus)
library(ggplot2)
library(gridExtra)
library(grid)
library(pander)
panderOptions('knitr.auto.asis', FALSE)
```


A function is created to analyze given distributions for data using the `fitdistrplus` library:
```{r func}
# extract ggplot legends: github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
g_leg <- function(a.gplot) {
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

find_fit <- function(x, dists) {
  # set up container for fits
  fits <- vector(mode = 'list', length = length(dists))
  names(fits) <- dists
  # fit each distribution
  for (i in 1:length(dists)) {
    d <- dists[i]
    fits[[d]] <- fitdist(x, d)
  }
  # create plots comparing fits
  p1 <- denscomp(fits, plotstyle = 'ggplot', legendtext = dists,
                 demp = TRUE, dempcol = 'black', fitlty = 1) +
    theme_minimal() + theme(legend.position='bottom', legend.direction = 'horizontal')
  p2 <- qqcomp(fits, plotstyle = 'ggplot', legendtext = dists, ylab = 'Emp. quantiles') +
    theme_minimal() + theme(legend.position='bottom', legend.direction = 'horizontal')
  p3 <- cdfcomp(fits, plotstyle = 'ggplot', legendtext = dists, fitlty = 1) +
    theme_minimal() + theme(legend.position='none')
  p4 <- ppcomp(fits, plotstyle = 'ggplot', legendtext = dists, ylab = 'Emp. probability') +
    theme_minimal() + theme(legend.position='none')
  # create legends
  leg_line <- g_leg(p1)
  leg_point <- g_leg(p2)
  # arrange plots & legends; draw box
  grid.arrange(p1 + theme(legend.position = 'none'), p2 + theme(legend.position = 'none'),
               p3, p4, leg_line, leg_point, nrow = 3, heights = c(0.43, 0.43, 0.14))
  grid.rect(width = 0.99, height = 0.99, gp = gpar(fill = NA))
  # get log likelihoods & find max
  l <- unlist(lapply(fits, `[[`, 'loglik'))
  f <- which.max(l)
  # create table of log likelihoods
  emphasize.strong.cols(f)
  pander(l)
  # get & return estimated parameters
  f <- names(f)
  e <- fits[[f]]$estimate
  return(e)
}
```

The function takes data and a vector of distributions as inputs and produces apprpriate plots and a table of log likelihood of the fit distributions, and returns parameter estimates for the fit with the best log likelihood.  The following libraries are required to execute this function:
```{r libraries-dummy, eval=FALSE}
library(fitdistrplus)
library(ggplot2)
library(gridExtra)
library(grid)
library(pander)
```

Data for problems 1-3 is combined into a single Excel file with multiple sheets and read into R:
```{r read-excel}
library(readxl)
q1 <- read_excel('wk8.xlsx', sheet = '06_01', col_names = FALSE)$X0
q2 <- read_excel('wk8.xlsx', sheet = '06_02', col_names = FALSE)$X0
q3 <- read_excel('wk8.xlsx', sheet = '06_03', col_names = FALSE)$X0
```

The distributions that will be fit to this data are stored:
```{r distributions}
d_cont <- c('norm', 'exp', 'lnorm', 'gamma', 'weibull')
d_disc <- c('geom', 'nbinom')
```


## Problem 1
The relevant plots and log likelihoods of the fit distributions are presented below:

```{r q1, results='asis', echo=FALSE}
e1 <- find_fit(q1, d_cont)
```

Based on the information above and the estimated fit paramters, the distribution that should be used to generate interarrival times is a log-normal distribution with parameters $\mu = `r round(e1[1], 4)`$ and $\sigma = `r round(e1[2], 4)`$.  The associated Simio expression is `Random.Lognormal(`r round(e1[1], 4)`, `r round(e1[2], 4)`)`.


## Problem 2
As above, the continuous distributions are fit to the data, returning the following results:

```{r q2, results = 'asis', echo=FALSE}
e2 <- find_fit(q2, d_cont)
```

Based on the information above and the estimated fit paramters, the distribution that should be used to generate call durations is a log-normal distribution with parameters $\mu = `r round(e2[1], 4)`$ and $\sigma = `r round(e2[2], 4)`$.  The associated Simio expression is `Random.Lognormal(`r round(e2[1], 4)`, `r round(e2[2], 4)`)`.


\newpage
## Problem 3
The same procedure as Problems 1 & 2 is followed, this time fitting discrete distributions:

```{r q3, results = 'asis', echo=FALSE}
e3 <- find_fit(q3, d_disc)
```

Based on the table above and the estimated fit paramters, the distribution that should be used to generate the number of extra tech-support people needed is a negative binomial distribution with parameters $r = `r round(e3[1], 4)`$ and $\mu = `r round(e3[2], 4)`$.

Simio accepts a probability parameter $p$ rather than a mean parameter $\mu$, where $p = r / (r + \mu)$; therefore the associated Simio expression is `Random.NegativeBinomial(`r as.character(round(e3[1], 0))`, `r e3[1]/(e3[1]+e3[2])`)`.


## Problem 4
The CDF for the uniform distribution is given by

$$
F_X(x) = 
\begin{cases} 
0 & x < a \\
\frac{x - a}{b - a} & a \leq x \leq b \\
1 & x > b
\end{cases}
$$
Setting the middle section equal to $u$ and solving for $u$ yields
$$F_X^{-1}(u) = u(b - a) + a$$



## Problem 5
The CDF for the Weibull distribution, for $x \geq 0$, is
$$F_X(x) = 1 - e^{(-x/\lambda)^k}$$
Setting the $F_X(x) = u$ and solving for $u$ yields
$$F_X^{-1}(u) = \lambda \left[ -\ln (1 - u) \right]^{1/k}$$



## Problem 8

```{r q8}
# set up provided info
produce <- c('oats', 'peas', 'beans', 'barley')
wholesale <- c(1.05, 3.17, 1.99, 0.95)
retail <- c(1.29, 3.76, 2.23, 1.65)
num_days <- 90

# set up demands
demand <- list()
demand[['oats']] <- data.frame(
  weight = c(0, 0.5, 1, 1.5, 2.0, 3.0, 4.0, 5.0, 7.5, 10.0),
  p = c(0.05, 0.07, 0.09, 0.11, 0.15, 0.25, 0.10, 0.09, 0.06, 0.03)
)
demand[['peas']] <- data.frame(
  weight = c(0, 0.5, 1.0, 1.5, 2.0, 3.0), p = c(0.1, 0.2, 0.2, 0.3, 0.1, 0.1)
)
demand[['beans']] <- data.frame(
  weight = c(0, 1.0, 3.0, 4.5), p = c(0.2, 0.4, 0.3, 0.1)
)
demand[['barley']] <- data.frame(
  weight = c(0, 0.5, 1.0, 3.5), p = c(0.2, 0.4, 0.3, 0.1)
)

# create container list for simulation results
walther <- list()

library(dplyr)
# create simulation for each list
for (p in 1:length(produce)) {
  # get 90 simulated sales for produce item
  set.seed(42) # set random number generator seed for replicability
  sales <- sample(demand[[produce[p]]]$weight, prob = demand[[produce[p]]]$p,
                  size = num_days, replace = TRUE)
  # store day number and calculate revenue, cost, and profit
  df <- data.frame(day_no = 1:num_days, sales) %>%
    mutate(revenue = sales * retail[p],
           cost = sales * wholesale[p],
           profit = revenue - cost)
  # add item name
  df$item <- rep(produce[p], num_days)
  # store in list
  walther[[p]] <- df
}

# collapse list into single data frame (tbl_df)
walther <- tbl_df(bind_rows(walther))

# calculate daily and running sums
daily <- walther %>% 
  group_by(day_no) %>% 
  summarise(revenue = sum(revenue),
            cost = sum(cost),
            profit = sum(profit)) %>%
  mutate(
    running_revenue = cumsum(revenue),
    running_cost = cumsum(cost),
    running_profit = cumsum(profit)
  )

# calculate total revenue, cost, and profit
tot <- daily[nrow(daily), 5:7]

# get daily and average sales by item
item <- walther %>% 
  group_by(item) %>% 
  mutate(avg_sales = cummean(sales))
```

\vspace{2em}

The simulation generated \$`r round(tot[1],2)` in revenue with an associated cost of \$`r round(tot[2], 2)`, yielding a profit of \$`r round(tot[3], 2)`.  The running totals of these numbers across the 90 days simulated are shown below, followed by plots of the daily demand for each item.

\newpage
```{r q8-plot, echo=FALSE}
ggplot(daily, aes(x = day_no)) +
  geom_line(aes(y = running_revenue, col = 'Revenue')) +
  geom_line(aes(y = running_cost, col = 'Cost')) +
  geom_line(aes(y = running_profit, col = 'Profit')) +
  scale_x_continuous('Day Number') + scale_y_continuous(NULL) +
  scale_color_discrete(NULL, breaks = c('Revenue', 'Cost', 'Profit')) +
  theme(legend.position = c(0.1, 0.9), legend.justification = c(0,1)) +
  labs(title = 'Running totals of revenue, profit, and cost',
       subtitle = 'Across all items over 90 days')

library(scales)
ggplot(item, aes(x = day_no)) +
  geom_col(aes(y = sales), alpha = 0.5) +
  geom_line(aes(y = avg_sales, col = item), show.legend = FALSE) +
  facet_wrap(~item, scales = 'free_y') +
  scale_y_continuous('Demand (lb)', breaks = pretty_breaks()) +
  scale_x_continuous('') +
  labs(title = 'Demand for each item sold',
       caption = 'Bars = Daily Value; Lines = Average Value')
```



